---
title: "Untitled"
output: html_document
date: "2025-02-25"
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}

library(bit64)
library(data.table)
library(ggplot2)
library(knitr)

data_folder    = "/Users/nichada/MyCode/MPCS/Job/_Interview/CBH"
```

```{r}
# Load the CSV file using read.csv()
shift_offers <- fread(file = paste0(data_folder, "/Shift Offers v3.csv"), stringsAsFactors = FALSE)

```

## Observe why some shift offers are not claimed

-   As a product team, I want to analyse how we can make our calim rate to be 100%. This will improve our users : workers and workplaces satisfactions, as well as drive revenue.

### Data cleaning & claimed column

```{r}
# Clean the canceled and deleted shifts
# create claimed column
shift_offers <- shift_offers[is.na(CANCELED_AT) & is.na(DELETED_AT)]
shift_offers[, c("CANCELED_AT", "DELETED_AT") := NULL]
shift_offers[, CLAIMED := ifelse(is.na(CLAIMED_AT), 0, 1)]

# Check the first few rows of unclaimed shifts
head(shift_offers)
```

### Claimed shifts

```{r}
# Filter for claimed shifts (where claimed_at is not NA)
claimed_shifts <- shift_offers[CLAIMED == 1]
  
# Optionally, if you want to exclude shifts that were later canceled or deleted:
unclaimed_shifts <- shift_offers[CLAIMED == 0]

# View the first few rows of claimed shifts
head(claimed_shifts)
```

-   Create a new variable Lead time hours

```{r}
# Create a new variable for lead time (in hours) between posting and shift start
shift_offers[, SHIFT_CREATED_AT := as.POSIXct(SHIFT_CREATED_AT)]
shift_offers[, SHIFT_START_AT := as.POSIXct(SHIFT_START_AT)]
shift_offers[, LEAD_TIME_HOUR := as.numeric(difftime(SHIFT_START_AT, SHIFT_CREATED_AT, units = "hours"))]
```

## Export csv to visualize

```{r}
# Export unclaimed shifts to a CSV file
fwrite(unclaimed_shifts, file = paste0(data_folder, "/unclaimed_shifts.csv"))

# Export claimed shifts to a CSV file
fwrite(claimed_shifts, file = paste0(data_folder, "/claimed_shifts.csv"))
```

## Analyze why unclaimed

summary - The unclaimed case is much greater than the claimed by more than 10x.

```{r}
cat("unclaimed counts: ", nrow(unclaimed_shifts), "\n")
cat("claimed counts: ", nrow(claimed_shifts))
```

```{r}
# Example: Compare shift slots
contingency_table <- table(shift_offers$CLAIMED, shift_offers$SLOT)
print(contingency_table)

# Example: Compare pay rates using boxplots
boxplot(unclaimed_shifts$PAY_RATE, claimed_shifts$PAY_RATE, 
        names = c("Unclaimed", "Claimed"), 
        main = "Pay Rate Comparison", ylab = "Hourly Rate")
```

### The Bad Workplace

### Method 1: Unclaimed Rate Method

-   What It Does: It calculates the raw percentage of unclaimed shifts per workplace by simply dividing the count of unclaimed shifts by the total shifts.

Strengths: - Straightforward and easy to understand.
- Directly shows the proportion of unclaimed shifts without any modeling assumptions.

Limitations: - Doesn’t account for variability in the number of shifts between workplaces.
- Doesn’t provide a statistical test for whether differences are due to chance.

```{r}
# Convert workplace_id to a factor, then to integer
shift_offers[, WORKPLACE_IDX := as.integer(as.factor(WORKPLACE_ID))]
shift_offers[, WORKPLACE_IDX := factor(WORKPLACE_IDX)]
```

```{r}
workplace_map <- unique(shift_offers[, .(WORKPLACE_IDX, WORKPLACE_ID)])
```

```{r}
# Group by workplace_id and calculate total shifts and unclaimed shifts per workplace
workplace_stats <- shift_offers[, .(
  total_shifts = .N,
  unclaimed_count = sum(CLAIMED == 0),
  claimed_count = sum(CLAIMED == 1)
), by = WORKPLACE_IDX]

# Calculate the unclaimed rate per workplace
workplace_stats[, unclaimed_rate := unclaimed_count / total_shifts]

# Order the results by the highest unclaimed rate
setorder(workplace_stats, -unclaimed_rate)

# View the results
print(workplace_stats)
```

```{r, fig.width=8, fig.height=20}
ggplot(workplace_stats, aes(x = reorder(WORKPLACE_IDX, unclaimed_rate), y = unclaimed_rate)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Unclaimed Shift Rate by Workplace", x = "Workplace", y = "Unclaimed Shift Rate")
```

### Method 2: Logistic Regression Method

-   What It Does: It models the probability of a shift being claimed (or unclaimed) for each workplace while testing if the differences are statistically significant.

-   How It Works: The regression estimates coefficients (in log-odds) for each workplace compared to a reference category.
    A negative and significant coefficient for a workplace means that, controlling for other factors (if any are included), shifts at that workplace are less likely to be claimed.
    The model provides p-values and confidence intervals, allowing you to assess which workplaces differ from the baseline in a statistically robust way.

-   Strengths: Provides formal statistical testing (p-values) to determine if a workplace’s unclaimed rate is significantly different from the reference.
    Can be expanded to include additional variables (like shift timing or pay rate) to control for confounding factors.

-   Limitations: If only workplace ID is included, it might largely reflect the raw differences; however, it still accounts for sample size variability and provides significance testing.
    The choice of reference category and dummy coding can influence interpretation.

```{r}
# Create a binary variable: 1 = Claimed, 0 = Unclaimed
# Run a logistic regression with workplace_id as a predictor.
# Note: This will create dummy variables for each workplace_id.
model <- glm(CLAIMED ~ WORKPLACE_IDX, data = shift_offers, family = binomial)
```

```{r}
model_summary <- summary(model)

# Convert the coefficients matrix into a data frame
coefs <- as.data.frame(model_summary$coefficients)

# Name the columns for clarity
# Typically, the columns are: Estimate, Std. Error, z value, and Pr(>|z|)
colnames(coefs) <- c("Estimate", "Std.Error", "z.value", "p.value")

# Create a column for the term names (row names)
coefs$Term <- rownames(coefs)

# (Optional) If you want to exclude the intercept from the sorting:
coefs <- coefs[coefs$Term != "(Intercept)", ]

# Sort by p-value in ascending order (lowest p-value = most significant)
coefs_sorted <- coefs[order(coefs$Estimate), ]
coefs_significant <- coefs_sorted[coefs_sorted$p.value < 0.05, ]
```

### Investigate the worst workplaces

Investigate the worst workplaces: The most negative, statistically significant workplaces (p \< 0.05) are likely causing unclaimed shifts.
Explore factors like pay rates, shift timing, or worker reviews at these locations.

```{r}
problematic_workplaces <- setDT(coefs_significant[coefs_significant$Estimate < 0, ])

setnames(problematic_workplaces, "Term", "WORKPLACE_IDX")
problematic_workplaces[, WORKPLACE_IDX := sub("^WORKPLACE_IDX", "", WORKPLACE_IDX)]

problematic_workplaces <- merge(
  problematic_workplaces,
  workplace_map,
  by = "WORKPLACE_IDX",   # common column
  all.x = TRUE            # left join; keep all rows from 'problematic_workplaces'
)

problematic_workplaces <- problematic_workplaces[, .(WORKPLACE_IDX, WORKPLACE_ID)]

kable(data.frame(problematic_workplaces), caption = "Problematic Workplaces - Coeficient < 0 ")
```

```{r}
# mark problematic workplace in the shift_offers data set
shift_offers[, problematic := factor(ifelse(WORKPLACE_IDX %in% problematic_workplaces$WORKPLACE_IDX, 1, 0))]
```

```{r}
# Compare pay_rate between the two groups
ggplot(shift_offers, aes(x = problematic, y = PAY_RATE, fill = problematic)) +
  geom_boxplot() +
  labs(
    title = "Pay Rate by Group",
    x = "Problematic Workplace",
    y = "Pay Rate"
  ) +
  theme_minimal()
```

```{r}
# Reshape (melt) multiple numeric columns
library(reshape2)  # or data.table::melt if you prefer
melted_data <- melt(
  data = shift_offers,
  id.vars = "problematic", 
  measure.vars = c("DURATION", "LEAD_TIME_HOUR")  # add more columns if needed
)

# Create boxplots for each variable in a facet
ggplot(melted_data, aes(x = problematic, y = value, fill = problematic)) +
  geom_boxplot() +
  facet_wrap(~ variable, scales = "free_y") +
  labs(
    title = "Comparison of Key Metrics by Group",
    x = "Problematic",
    y = "Value"
  ) +
  theme_minimal()
```

```{r}
# Compute summary statistics for each group
shift_offers[, .(
  count = .N,
  mean = mean(LEAD_TIME_HOUR, na.rm = TRUE),
  median = median(LEAD_TIME_HOUR, na.rm = TRUE),
  sd = sd(LEAD_TIME_HOUR, na.rm = TRUE),
  min = min(LEAD_TIME_HOUR, na.rm = TRUE),
  max = max(LEAD_TIME_HOUR, na.rm = TRUE),
  q25 = quantile(LEAD_TIME_HOUR, 0.25, na.rm = TRUE),
  q75 = quantile(LEAD_TIME_HOUR, 0.75, na.rm = TRUE)
), by = problematic]
```

```{r}
ggplot(shift_offers, aes(x = PAY_RATE, fill = problematic)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Pay Rate Distribution",
    x = "Pay Rate",
    y = "Density"
  ) +
  theme_minimal()
```

```{r}
ggplot(shift_offers, aes(x = LEAD_TIME_HOUR, fill = problematic)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Lead Time Hour Distribution",
    x = "Pay Rate",
    y = "Density"
  ) +
  theme_minimal()
```

```{r}
ggplot(shift_offers, aes(x = factor(problematic), fill = SLOT)) +
  geom_bar(position = "fill") +
  labs(
    title = "Slot Distribution by Group",
    x = "Slot",
    fill = "Problematic (0 = Non-Problematic, 1 = Problematic)"
  ) +
  theme_minimal()
```

Problematic vs Non-Problematic workplace characteristics : 1.
There is not a significant impact in Pay rate 2.
There is not a significant impact in Duration 3..
Problematic workplace has shorter lead time than the non-problematic ones.
- Comparing median 22.63 for non-problematic and 16.07 for problematic.
- Problematic workplaces have narrower interquartile range.
Q25 to Q75 for Problematic = 4.89 to 76.00 (range \~71).
Q25 to Q75 for Non-Problematic = 7.05 to 193.87 (range \~186).
- Negative Lead Time : Non-Problematic min = -870.93 hours (suggesting potential data errors or shifts posted after they started).
Could indicate shifts that were backfilled.
4.
Problematic workplace has more PM than AM slots than the non-problematic ones.
However the data doesn't show a significant sign.
5.
There could be a problem with the workplace name itself that is not desirable.

\newpage

## Clean data2

Colinearity

```{r}
unique(shift_offers$CHARGE_RATE)
var(shift_offers$CHARGE_RATE, na.rm = TRUE)
var(shift_offers$PAY_RATE, na.rm = TRUE)
cor(shift_offers$CHARGE_RATE, shift_offers$PAY_RATE, use = "complete.obs")
```

## Expand Models

-   Explore other factors that are contributing to the unclaimed offers
-   Use log regression

```{r}
model_full <- glm(CLAIMED ~ WORKPLACE_ID + PAY_RATE + DURATION + SLOT + LEAD_TIME_HOUR, 
                  data = shift_offers, family = binomial)
summary(model_full)
```

-   Separate lines for PAY_RATE, CHARGE_RATE, DURATION, etc. Each of these has its own coefficient and p-value, indicating how they affect the odds of a shift being claimed after controlling for workplace and other variables.
-   Note include CHARGE_RATE due because it may not be a factor of consideration for worker.

model_full \<- glm(CLAIMED \~ WORKPLACE_ID + PAY_RATE + CHARGE_RATE + DURATION + SLOT + LEAD_TIME_HOUR, data = shift_offers, family = binomial)

### Find Cut-out points

```{r}
# Create a new data table spanning the range of rate values
new_data <- data.table(
  PAY_RATE = seq(min(shift_offers$PAY_RATE), max(shift_offers$PAY_RATE), length.out = 100),
  DURATION = median(shift_offers$DURATION),
  SLOT = factor("am", levels = unique(shift_offers$SLOT)),  # Ensure it matches existing levels
  LEAD_TIME_HOUR = median(shift_offers$LEAD_TIME_HOUR, na.rm = TRUE),
  WORKPLACE_ID = factor("64316d145665f001ba8ff5f9", levels = unique(shift_offers$WORKPLACE_ID))
)

# Compute the predicted probabilities using the expanded model
pred_probs <- predict(model_full, newdata = new_data, type = "response")

# Plot predicted probability vs. rate
plot(new_data$PAY_RATE, pred_probs, type = "l", xlab = "Pay Rate", ylab = "Predicted Probability of Being Claimed", ylim = c(0, 1))
abline(h = 0.5, col = "red", lty = 2)  # a reference line for 50% probability

```

```{r}
# Create a new data table spanning the range of rate values
new_data <- data.table(
  LEAD_TIME_HOUR = seq(min(shift_offers$LEAD_TIME_HOUR), max(shift_offers$LEAD_TIME_HOUR), length.out = 100),
  PAY_RATE = median(shift_offers$PAY_RATE, na.rm = TRUE),
  DURATION = median(shift_offers$DURATION),
  SLOT = factor("am", levels = unique(shift_offers$SLOT)),  # Ensure it matches existing levels
  WORKPLACE_ID = factor("64316d145665f001ba8ff5f9", levels = unique(shift_offers$WORKPLACE_ID))
)

# Compute the predicted probabilities using the expanded model
pred_probs <- predict(model_full, newdata = new_data, type = "response")

# Plot predicted probability vs. rate
plot(new_data$LEAD_TIME_HOUR, pred_probs, type = "l", xlab = "LEAD_TIME_HOUR", ylab = "Predicted Probability of Being Claimed", ylim = c(0, 1))
abline(h = 0.5, col = "red", lty = 2)  # a reference line for 50% probability

```

## Conclusion

-   The analysis indicates that workplace is a key factor driving unclaimed shift offers.
-   I further investigated other potential reasons why certain workplaces are less desirable, including pay rate, shift duration, time slot, and lead time.
-   Shorter lead times appear to contribute to the issue, particularly in problematic workplaces. However, since some non-problematic workplaces also have short lead times, this factor alone is not necessarily a deal-breaker. Instead, maintaining a reasonable lead time is likely beneficial but not the sole determinant of shift claims.
-   The most concerning finding is that some workplaces have an extremely high unclaimed rate (\>90%), as evidenced in workplace_stats. This suggests a need for further investigation into the work environment at these **locations**, including **workplace conditions**, **location**, and **type of work**.
-   Understanding these factors will help us determine why these workplaces struggle to get shifts claimed and identify potential interventions to improve their claim rates.

## Recommendation to the Product Team

### Short Term Actions

### Long Term Strategy

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
